### 2011-04-29 RemovedText.tex

Open workflow systems such as Taverna \cite{Hull:2006p60} and KNIME\cite{WebKNIME}
have opened the door for 
an easy and seamless integration of BO components into executable workflows 
or data pipelines. A combination of the Chemistry Development Kit and Taverna, for instance, was
reported in 2010 \cite{Kuhn:2010p4001}.

KNIME currently has a built-in collection of CDK-based and
OpenBabel-based nodes\cite{WebKNIMEExtensions} for implementing
chemistry workflows. Another collection of chemistry nodes---based on
the RDKit and its Java wrapper---appeared in 2010 and is available to
KNIME users from the KNIME's official ``Community Updates'' site.
Indigo nodes for KNIME are in preparation and will appear nearby the
RDKit nodes by May 2011.